\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-06-15 paper/paper.tex
normal and special cases

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

%%% Nach Lehrbuch müssten die methodischen Schritte hier beschrieben werden.

% Defining income
% Beschreiben, dass sich aus den publizierten Einkommensgrössen und den Steuerbeträgen unterschiedliche 
% Einkommen berechnen lassen
% Es steht der Vorwurf im Raum, die unteren Einkommen seien falsch abgebildet, weil die Sozialhilfe nicht versteuert ist

@@

%%
% Population coverage
% Analysen zu den Nullern aufzeigen> ist eine Hauptkritik an der Nutzung von Steuerdaten
% Die Bedeutung von Normal- und Sonderfällen
% Den Vergleich mit der HABE beschreiben, welche Einkommensgrösse nutzen wir aus der HABE um diese möglichst mit den Steuerdaten vergleichbar zu machen? Wir brauchen das Bruttoeinkommen nach Abzug der Sozialversicherungsbeiträge, dies lässt sich mit dem Reineinkommen vergleichen

@@

\subsection{Zero share}

Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later)

\subsection{Income measures with and without zeros}

>>>>

\begin{document}

@@

%%% Nach Lehrbuch müssten die methodischen Schritte hier beschrieben werden.
% Man könnte zu den Zeitreihen jeweils die Varianz ausweisen-> dass würde helfen die Reihe zu beurteilen. Man sieht dann: aha, die Reihe ist im Vergleich zur anderen höher/tiefer. Ob die Reihen stärkeren Schwankungen unterliegen könnte einfach aus der Varianz gelesen werden

% Defining income
% Beschreiben, dass sich aus den publizierten Einkommensgrössen und den Steuerbeträgen unterschiedliche 
% Einkommen berechnen lassen
% Es steht der Vorwurf im Raum, die unteren Einkommen seien mit Steuerstatistiken schlecht abgebildet, weil bedarfsabhängige Leistungen nicht versteuert sind

% Es müsste eine Reihe zum gesetzliche Grundbedarf vorliegen. Wüsste ich jetzt grad nicht von wo...

@@

% Calculate the gini
% Ein bisschen erklären, wie aus gruppierten Tabellen der Gini-Koeffizient berechnet wird.

% Polaritätsindex
% Alderson and Doran (2013:56) beschreiben zwei mögliche Indizies Median relative polarization index (MRP) eingeführt von (Morris, Bernhardt und Handcock 1994:217), der in den lower relative polarization index(LRP) und den upper relative polarization index (URP) zerlegt werden kann (ebd.209)
% (1. Schritt) Probability density functions
% (2. Schritt) Overall relative density function -> Achtung diese Dichtefunktion bildet sowohl die  Lokationsverschiebung als auch durch Verschiebung der Form ab. Eigentlich interessiert aber, der shape shift
% Deshalb wird die Verteilung um die Lokationsverschiebung bereinigt und nur der shape-shift angeschaut. Ausgewiesen werden (MRP)-> hat polarisierung stattgefunden: LRP und URP-> ist polarisierung eher auf eine Zunahme armer (LRP) zurückzuführen oder auf eine Zunahme von Reichen? (URP)

%%
% Population coverage
% Analysen zu den Nullern > ist eine Hauptkritik an der Nutzung von Steuerdaten
% Die Bedeutung von Normal- und Sonderfällen
% Den Vergleich mit der HABE beschreiben, welche Einkommensgrösse nutzen wir aus der HABE um diese möglichst mit den Steuerdaten vergleichbar zu machen? Wir brauchen das Bruttoeinkommen nach Abzug der Sozialversicherungsbeiträge, dies lässt sich mit dem Reineinkommen vergleichen

@@

\emph{Gini coefficients considering mean-tested benefits}

\subsection{Measuring inequality}

% Zero-shares (alter Textbaustein)
%Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later) 
% Das allenfalls in der Methodensection machen?
% Vergleich mit der Dell-Reihe zu Non-fillers
% Überschneidungen gib es leider nicht 
% Interessant wäre es auch, zu sehen ob es Brüche in den Zeitreihen der Zahl der Steuerpflichtigen und jener ohne Besteuerung gibt.

\subsection{Population coverage}

\emph{Normal versus special cases}

The FTA stopped to publicly report data for special cases after tax period 1993/94. 
For more recent periods however we can compare data from the 
BRUELHARTPROJEKT (wie nennen wir das?) to the FTA normal cases, as these data are 
identical but the BRÜLHART data include special cases. 
We will have a look at both, the 1993/94 period as the last period where both 
numbers were reported as well as 2009 where we compare data from the same source 
but using differnt data sheets.

%ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]

\includegraphics[width=\maxwidth]{figure/specialcases93941} \caption[]{\label{fig:specialcases93941}}
\end{figure}

\begin{figure}[]

\includegraphics[width=\maxwidth]{figure/specialcases93942} \caption[]{\label{fig:specialcases93942}}
\end{figure}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]

\includegraphics[width=\maxwidth]{figure/specialcases091} \caption[]{\label{fig:specialcases091}}
\end{figure}

\begin{figure}[]

\includegraphics[width=\maxwidth]{figure/specialcases092} \caption[]{\label{fig:specialcases092}}
\end{figure}

As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
differnt from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
million tax units defined as being normal cases compared to 0.51 million special cases.}
In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
within all cases compared to normal cases only. The share of special cases however decreased 
to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}

\emph{Income measures with and without zeros}

### 2014-06-15 paper/subdocuments/results.Rnw
normal and special cases

>>>>

\emph{Normal versus special cases}

The FTA stopped to publicly report data for special cases after tax period 1993/94. 
For more recent periods however we can compare data from the 
BRUELHARTPROJEKT (wie nennen wir das?) to the FTA normal cases, as these data are 
identical but the BRÜLHART data include special cases. 
We will have a look at both, the 1993/94 period as the last period where both 
numbers were reported as well as 2009 where we compare data from the same source 
but using differnt data sheets.

%ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
<<specialcases9394, echo=FALSE,warning=FALSE,fig.cap=''>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(ggplot2)
#library(reldist)
normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
#special <- read.csv("data/ginis_und_perzentile_sonder.csv",sep="\t")
normal$case <- "normal case"
#special$case <- "special case"
pooled$case <- "pooled, normal and special cases"
d<-rbind(normal,pooled)
d<-d %.% 
  filter(kanton=="CH",steuerperiode==1993.5)

start <- which(names(d)=="p1")
end <- which(names(d)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(d[1,start:end]),percentile,case="normal")
pooled <- data.frame(t(d[2,start:end]),percentile,case="pooled")
names(normal)[1]=c("inc")
names(pooled)[1]=c("inc")
percentiles <- rbind(normal,pooled)  
ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
	geom_line()+theme_bw()+
	scale_colour_grey()+
	xlab("Percentile")+ylab("taxable income")+
	ggtitle("Taxable incomes of normal and special cases 1993/94")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))

#plot relative distribution
rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
rel$change <- rel$pooled.inc/rel$normal.inc
ggplot(rel, aes(x=pooled.percentile,y=change))+
	geom_line()+theme_bw()+
	scale_colour_grey()+
	geom_line(linetype="dashed",aes(y=1),size=2)+
	xlab("Percentile")+ylab("taxable income")+
	ggtitle("Taxable incomes of normal versus special cases 1993/94")

#steuerpflichtige estv / stpf bruelhart
#1- (2762419 / 3268548)
@

<<specialcases09, echo=FALSE,warning=FALSE,fig.cap=''>>=
library(foreign)
library(ggplot2)
#library(reldist)
normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2009)
start <- which(names(normal)=="p1")
end <- which(names(normal)=="p99")
percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
normal <- data.frame(t(normal[1,start:end]),percentile)
names(normal)[1]=c("inc")

#bruelhart data
pooled <- filter(read.dta("../../data/data_Schweiz.dta"), Veranlagungsperiode==2009, Einheit=="Total")
start <- which(names(pooled)=="p1")
p95 <- which(names(pooled)=="p95")
end <- which(names(pooled)=="p99")
pooled <- data.frame(t(pooled[1,c(start:p95,end)]),percentile)
names(pooled)[1]=c("inc")
pooled$inc<-pooled$inc/1000 #adjust scale
normal$case <- "normal case"
pooled$case <- "pooled, normal and special cases"
percentiles <- rbind(normal,pooled)
ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
	geom_line()+theme_bw()+
	scale_colour_grey()+
	xlab("Percentile")+ylab("taxable income")+
	ggtitle("Taxable incomes of normal and special cases 2009")+
  theme(legend.justification=c(0,1),legend.position=c(0, 1))

#plot relative distribution
rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
rel$change <- rel$pooled.inc/rel$normal.inc
ggplot(rel, aes(x=pooled.percentile,y=change))+
	geom_line()+theme_bw()+
	geom_line(linetype="dashed",aes(y=1),size=2)+
	scale_colour_grey()+
	xlab("Percentile")+ylab("taxable income")+
	ggtitle("Taxable incomes of normal versus special cases 2009")

#steuerpflichtige estv / stpf bruelhart
#1- (3422210 / 3689063)
@

As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
differnt from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
million tax units defined as being normal cases compared to 0.51 million special cases.}
In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
within all cases compared to normal cases only. The share of special cases however decreased 
to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}
